Open Preprints in Ecology \& Evolution

### 2012-09-10 ms.tex
started adding Karthik's notes to the ms

\author{Joel J. Adamson}
\affiliation{Ecology, Evolution and Organismic Biology, University of North Carolina at Chapel Hill}

@@

@@

@@

Preprint servers also establish priority in a fair way. Some manuscript will
spend much more time in the review process. Public preprints server offer a
much fairer way to establish intellectual priority by making the work
available when done, and even if the exact organisation of the manuscript may
change. Surprisingly, there is perception in biology that public preprints
make it easier to steal ideas, as if scientific ideas only took form in
published material.
% TP This needs more development, I'm not sure I get your point
Mathematicians and physicists have embraced arXiv in part
to establish priority in a fair way\cite{cal12}.

Some of the responses to public preprints are surprising since they are,
essentially, the same as exchanging preprints among colleagues.
% TP Well, no. You trust your colleagues, and you trust that they will not give your
%    MS to anyone!
Prepublication reviews by a small network of colleagues is an important part
of the scientific process, which is attested by the fact that nearly all
published papers acknowledge comments by people not listed as co-authors.
Preprints servers simply offer a way to extend this network of colleagues to
the entire scientific community. It ensures that science is not constrained by
small networks of scientists exchanging ideas. Ginsparg made arXiv.org in part
for democratic reasons: he wanted everyone from graduate students in small
universities to Princeton professors to have access to the most recent
scientific \emph{ideas}.
% TP One important point to make: the MS and the idea are different things.
%    We should focus more on the later...
Ginsparg revolutionary idea was simply to use the power of the internet for
preprints, not just for the end product, so the process can be open from A to
Z, instead of being just open at the end of the process.
% TP Since when is the publishing buisness an open process? :-)

@@

While the practice is still rare, preprints are becoming more common in
biological sciences, which is experiencing faster growth in arXiv submissions
than any other fields \cite{cal12}. Also, most scientific journals are
preprint-friendly: Nature, PLOS, BMC, PNAS, Science (mostly)
\ref{table:policies}, and all the journals from Elsevier and Springer. Very
recently, the Ecological Society of America recently changed its policy to allow
public preprints (REF). In our field, few scientific publications will not
consider a manuscript submitted to arXiv. Still, many ecology \& evolution journals
adopt a ``by default'' hostile attitude towards preprints, mostly due to the
lack of clear policy of the publishers. As an example, Wiley-Blackwell, which
publishes some of the leading journal in the field, has no official policy on
the subject \ref{table:policies}.

\begin{table*}
    \centering
    \begin{tabular}{|ll|}
    \hline
    Publisher                                   & Policy \\
    \hline
    Springer                            	& Accept \\
    BMC                                 	& Accept \\
    Elsevier                            	& Accept \\
    Nature Publishing Group             	& Accept \\
    Public Library of Science           	& Accept \\
    Royal Society                       	& Accept \\
    National Academy of Science (USA)           & Accept \\
    Science                             	& Accept/Ambiguous \\
    Wiley-Blackwell                       	& No general policy \\
    Ecological Society of America       	& Refuse \\
    British Ecological Society                  & ? \\
    \hline
    \end{tabular}
    \caption{Policies for important publishers in ecology and evolution.}
    \label{table:policies}
\end{table*}

There are other ``cultural'' barriers to adoption of preprint servers
by ecologists and evolutionary biologists.  The most notable preprint
server, arXiv, is designed to work best with \TeX{}, a nearly
universal document preparation system among physicists and
mathematicians.  Jackson \cite{jackson2002preprints} further argues
that \TeX{} introduced an ``open source mindset'' to its users, who
now freely share their research findings as well as their software.
Many biologists instead use proprietary software to prepare their
research findings, and many ecology and evolution journals officially
prefer Microsoft Word documents as submissions.  The recent
discipline-wide adoption of free software packages such as R, as well
as interest in open access publishing, has coincided with recent use
of preprint servers by biologists, supporting Jackson's claims.

\section{Current offerings}

@@

% @Joel
% Rough outline:
% 1. Description of arXiv (origins, funding, maintenance)
% 2. Submission process
% 3. Role within math and physics community (nearly universal), including recent changes
% 4. Nature of citations and "priority"
%
% My biggest problem here is how to cite my sources: the information
% on funding comes from a press release from Cornell University
% Libraries; other basic "raw data" comes straight from the arXiv
% webpages. How do I cite that?

arXiv (\url{http://arxiv.org/}) is the most widely-used preprint
server today, and its use is almost universal in mathematics and most
branches of physics.  Physicist Paul Ginsparg originated arXiv in 1991
for theoretical high-energy physicists to communicate preprints via
email and ftp, and soon thereafter adopted the newly created
world-wide web\cite{jackson2002preprints}.  arXiv now receives over
7,000 submissions per month
(\url{http://arxiv.org/show_monthly_submissions}).  arXiv divides its
submissions into subcategories of physics, mathematics, computer
science, quantitative biology, finance and statistics.  The
quantitative biology category includes subcategories for Populations
and Evolution, Quantitative Methods and other categories that may be
of interest to ecologists and evolutionary biologists.
% Check me on this @Pierre!  Not only is this vague, but I'm not sure
% how to verify it; I want to give people the impression that a
% "Quantitative ecology" or "General ecology" category could be
% created easily
The creation of new categories is regulated by the same system that
categorizes papers.

Submission to arXiv is fully automated via the world-wide web.
Authors can submit \TeX{}/\LaTeX{} documents or directly in PDF or
PostScript (for example, as exported by a word processor).  A
moderation system was put in place in 2004: papers must be categorized
by an ``endorser.''  At least one author of a paper must be an
endorser that has previously submitted a paper or has received
permission to submit to a particular category.  Many authors in
mathematics and physics submit papers as soon as they are ready for
review by colleagues, although another popular option is submitting
simultaneously to a journal and arXiv.

Most papers posted to arXiv are eventually printed in journals.  There
are notable exceptions, including Perelman's landmark paper leading to
the proof of the Poincar\'{e} conjecture \cite{2002math.....11159P}.
arXiv provides a reliable citation system for all eprints (see our
citation \cite{2002math.....11159P}), lending a form of ``intellectual
priority'' to works posted there.  Despite these marks of arXiv 
functioning as a scientific journal, arXiv has never sought to replace
scientific journals and explicitly states that it serves a different
function as ``an openly accessible, moderated repository for scholarly
articles in specific scientific disciplines.''

arXiv is now administered by the Cornell University Libraries.
Funding comes from voluntary pledges by academic institutions along
with matching funds from the Simons Foundation \cite{arxiv_future}.
One-hundred twenty six of the top two-hundred institutions in terms of
downloads have provided the operating budget for arXiv over the next
five years.  This plan reduces the financial burden on Cornell
University and transfers governance to a collaborative community in
accordance with arXiv's key principles.  arXiv takes numerous measures
to ensure that the repository will remain permanently available and
submissions will be readable.

>>>>

@@

\author{Karthik Ram}
\affiliation{Environmental Science, Policy, and Management. University of California, Berkeley. Berkeley, CA. United-States of America.}

@@

\section{Introduction}

Open preprints servers allow authors to make their manuscripts publicly available before, or in parallel to, submitting them to journals for traditional
peer-review. This idea gained popularity 20 years ago with the advent of arXiv, an open preprint server, widely used in the physical sciences \cite{gin11}. The idea behind pre-print servers is fundamentally simple: to make the results of a scientific endavour available to the scientific community as soon as possible rather than wait until the peer-review process is fully completed. The point of arXiv and open preprints servers is not to avoid the peer-review. Almost all manuscripts submitted to arXiv undergo formal peer-review at a journal. The point is to open an important phase of the publication cycle ...

In this article, we will first highlight the advantages of open preprints
servers for both scientists and publishers. We will also debunk a few
misconceptions, discuss the policies of major publishers in ecology an
evolution, and briefly review the most popular open preprint servers.

@@

% @to myself: rewrite this paragraph:
Some of the responses to public preprints are surprising since they are,
essentially, the same as exchanging preprints among colleagues.
% TP Well, no. You trust your colleagues, and you trust that they will not give your
%    MS to anyone!
Prepublication reviews by a small network of colleagues is an important part
of the scientific process, which is attested by the fact that nearly all
published papers acknowledge comments by people not listed as co-authors.
Preprints servers simply offer a way to extend this network of colleagues to
the entire scientific community. It ensures that science is not constrained by
small networks of scientists exchanging ideas. Ginsparg made arXiv.org in part
for democratic reasons: he wanted everyone from graduate students in small
universities to Princeton professors to have access to the most recent
scientific \emph{ideas}.
% TP One important point to make: the MS and the idea are different things.
%    We should focus more on the later...
Ginsparg revolutionary idea was simply to use the power of the internet for
preprints, not just for the end product, so the process can be open from A to
Z, instead of being just open at the end of the process (and often not open at
all).
% TP Since when is the publishing buisness an open process? :-)
% PhDP: Wishful thinking :) Corrected...

Preprint servers also establish priority in a fair way. Since some manuscripts
will spend much more time in the review process, public preprints servers offer
a fairer way to establish intellectual priority by making the work available
when done. Surprisingly, there is perception in biology that public preprints
make it easier to steal ideas, while mathematicians and physicists have embraced
arXiv in part to establish intellectual priority in a fair way \cite{cal12}.

Preprint servers also provide additional opportunities for review, often from members outside the list of authors or their immediate peers (cite). This process can further improve the quality of a manuscript before it reaches potential reviewers. pre-prints servers can therefore reduce reviewer load/burden by opening up oppportunities for additional feedback.

% Karthik's section
% I'll need to flesh this section out a bit more but here are the two main points
% pre-print servers enrich the review process by providing a venue to solicit comments from anyone. Manuscripts can undergo several rounds of revision during this stage, resulting in a higher quality manuscrpt reaching reviewers than ones that have not undergone this process. There are various benefits here. First, it can reduce the time burden on reviewers, the need for multile rounds of revision.

% Second point, pre-publication provides a scaffolding the possibility of a implementing a reward system for reviewers and publishers. Open reviews can be judged on their quality and usefulness both by the authors and by other members of the community. Reviews and comments deemed useful by the community can be used to reward contributers with reputation, which then provides a measure of trustworthiness. The reputation system can also serve to deter malicious reviews and personal attacks, and such penalties can limit  @PhDP: Do you think this is a useful direction to go? It could be moved into the discussion or stay with the benefits of pre-print servers.

@@

While submitting to public preprints servers is still uncommon in ecology and
evolution, preprints are becoming more common in biological sciences. The
quantitative biology section in arXiv is experiencing faster growth in
submissions than any other fields \cite{cal12}. Also, most scientific journals
are preprint-friendly: Nature, PLOS, BMC, PNAS, Science (mostly)
\ref{table:policies}, and all the journals from Elsevier and Springer. Very
recently, the Ecological Society of America amended its policy to allow
public preprints (REF). In our field, few scientific publications will not
consider a manuscript submitted to arXiv.  Still, many ecology \& evolution
journals adopt a ``by default'' hostile attitude towards preprints, mostly due
to the lack of clear policy of the publishers. As an example, Wiley-Blackwell,
which publishes some of the leading journal in the field, has no official policy
on the subject \ref{table:policies}.

\begin{table*}
    \centering
    \begin{tabular}{|ll|}
    \hline
    Publisher                                   & Policy \\
    \hline
    Springer                            	& Accept \\
    BMC                                 	& Accept \\
    Elsevier                            	& Accept \\
    Nature Publishing Group             	& Accept \\
    Public Library of Science           	& Accept \\
    Royal Society                       	& Accept \\
    National Academy of Science (USA)           & Accept \\
    Ecological Society of America       	& Accept \\ % Victory!
    Science                             	& Accept/Ambiguous \\
    Wiley-Blackwell                       	& No general policy \\ % grrrrrr
    British Ecological Society                  & ? \\ % Waiting for their answer...
    % Oxford worth mentionning?
    \hline
    \end{tabular}
    \caption{Policies for important publishers in ecology and evolution.}
    \label{table:policies}
\end{table*}

\section{Current offer}

@@

arXiv (\href{http://arxiv.org/}{http://arxiv.org/}).
arXiv is funded by a network of universities.

% @Joel
